---
title: "Prob1_MA568"
author: "Hengchang Hu"
date: "9/29/18"
output:
  html_document: default
  pdf_document: default
---

#### 1.Download data file Retinal_ISIs.txt which contains wating times in milliseconds & Plot spiking activity as a histogram of the distribution of the times  and as a spike train time series  & Describe the spiking properties .  


```{r histgram, echo=FALSE}
Retinal_ISIs <- read.table("C:/Users/hugo1/Documents/MA568/Retinal_ISIs.txt")

waitingTime <- as.numeric(Retinal_ISIs[,1])

hist(waitingTime, breaks = seq(0,200,1), xlab = 'wating time (ms)', ylab = 'Counts', main = "Hisogram of the distribution of times", col = 'red') # hisogram of the distribution of times

arriveTime <- rep(0,972)

arriveTime[1] <- waitingTime[1]
for (i in 1:971)
{
  arriveTime[i + 1] <- arriveTime[i] + waitingTime[i + 1]
}

hist(arriveTime, breaks = 1:30000,xlab = "Arrive Time (ms)", ylab = NULL, main = "Spike train time series") # spike train time series
```

#### **Conclusion: The histogram shows that most of the waiting times are 3-30 ms, and the waiting times between 3-11 ms consists of more than 50 % of observed data. So a homogeneous Poisson Process may not well fitted in this data**.

* * *

#### 2.Compute a 5-number summary & box plot for ISI distribution  


```{r Statistics, echo=FALSE}
summary(waitingTime) # get min, .25 quantile, median, .75 quantile & max

boxplot(waitingTime) # box plot for ISI distribution

```

#### **Conclusion: 5-number summary of this data is (2, 5, 10, 44, 190) corresponding to (min, .25 quantile, median, .75 quantile, max)**.
* From the 5-number summary, we can see that most of waiting time data are concentrated in 5-44 ms.
* From the boxplot, we can tell that this is a long-tailed distribution because there are lots of big data over 100 ms although most of data are concentrated in 5-44 ms.


* * * 

#### 3.Bin the spike train data from Retinal_ISIs.txt into time bins of width 1ms., 10ms., 100ms & Plot time series of spike counts and distribution of spike counts as histogram for each bin width  


```{r echo=FALSE}
hist(arriveTime, breaks = 1:30000, xlab = 'Arrive Time (ms)', ylab = 'Counts', main = "Time series of Spike counts for 1ms width time bins") # time series of spike counts as a histogram for time bins of width 1ms

hist(c(rep(1,972), rep(0,30000-972)), breaks = seq(0,1,0.5), main = "Histogram of distribution of spike counts for 1ms width time bins", xlab = "spike counts", col = 'red') # distribution of spike counts as a histogram for time bins of width 1ms

hist(arriveTime, breaks = seq(0,30000,10), xlab = 'Arrive Time (ms)', ylab = 'counts', main = "Time series of Spike counts for 10ms width time bins") # time series of spike counts as a histogram for time bins of width 10ms

hist(c(rep(0,2240), rep(1,574), rep(2,160), rep(3,26)), main = "Histogram of distribution of spike counts for 10ms width time bins", xlab = "spike counts", col = 'red') # distribution of spike counts as a histogram for time bins of width 10ms

hist(arriveTime, breaks = seq(0,30000,100), xlab = 'Arrive Time (ms)', ylab = 'counts', main = "Time series of Spike counts for 100ms width time bins") # time series of spike counts as a histogram for time bins of width 100ms

hist(c(rep(0,22), rep(1,45), rep(2,59), rep(3,61), rep(4,37), rep(5,26), rep(6,23), rep(7,14), rep(8,8), rep(9,3), rep(10,1), rep(11,1)), breaks = seq(0,11,0.5), main = "Histogram of distribution of spike counts for 100ms width time bins", xlab = "spike counts", col = 'red') # distribution of spike counts as a histogram for time bins of width 100ms
```

#### **Conclusion: Comparing to model for increments of a Poisson process, the distribution of spike counts are long-tailed especially in 100 ms width time bins**.

* * *  

#### 4.Plot the likelihood as a function of $\lambda$ for values of $\lambda$ between 0 Hz to 60 Hz & Find value $\hat\lambda_{ML}$ that maximize the likelihood & Provide an approximate 95% confidence interval for $\hat\lambda_{ML}$.  


```{r, echo = FALSE}
loglikelihood <- function(x) 972 * log(x) - 30 * x # create a log likelihood function

plot(loglikelihood,from = 0, to = 60, col = 'red') # plot this function for values of lambda between 0Hz to 60Hz

optimize(loglikelihood, lower = 0, upper = 60, maximum = T) # get maximum value of log likelihood function
```

#### **Conclusion: The value of $\hat\lambda_{ML}$ that maximaize the likelihood is 32.4, and the approximate 95% confidence interval for $\hat\lambda_{ML}$ is $[32.4 - 1.96 \times \frac{\sqrt{972}}{30}, 32.4 + 1.96 \times \frac{\sqrt{972}}{30}]$, i.e. $[30.36311, 34.43689].$**

* * *  

#### 5.Plot an empirical CDF of the interspike intervals for the data & Plot the exponential CDF on the same plot as your empirical CDF & Construct a KS plot of the empirical CDF on the x-axis against the model CDF on the y-axis

```{r, echo = FALSE}
expCdfofWaitingTime <- function(x) 1 - exp(x / -32.4) # CDF of exponential distribution of waiting time

eCDFofWaitingTime <- ecdf(waitingTime) # empirical CDF of waiting time

plot(eCDFofWaitingTime(0:200), xlab = "Waiting Time (ms)", ylab = " ", main = "Empirical Distributions", type = "l", lty = 2) # plot the empirical cdf of waiting time

plot(expCdfofWaitingTime,0,200,ylab = " ",col = 'red', add = TRUE) # add exponential CDF to second plot

legend(100,0.4, legend = c("Empirical Distribution", "Model Distribtuion"), col = c("black", "red"), lty = 2:1)

ecdfPoint <- eCDFofWaitingTime(0:200) # get ecdf value for corresponding x

mcdfPoint <- expCdfofWaitingTime(0:200) # get model cdf value for corresponding x

diagLine <- function(x) x # create diag line function 

lowerConfidenceInterval <- function(x) x - (1.36 / sqrt(972)) # confidence bounds

upperConfidenceInterval <- function(x) x + (1.36 / sqrt(972))

plot(x = ecdfPoint, y = mcdfPoint, type = "l", xlab = "Empirical CDF", ylab = "Model CDF", main = "KS plot", col = 'red') # KS plot

plot(diagLine,0,1,add=TRUE)

plot(lowerConfidenceInterval, 0,1,lty = 2,add=TRUE)

plot(upperConfidenceInterval, 0,1,lty = 2,add=TRUE)

KSstatistics <- max(abs(mcdfPoint - ecdfPoint))

print(paste('KS statistics is', KSstatistics))

```


* * *  

#### 6.Construct a QQ plot of the empirical vs model quantiles


```{r, echo = FALSE}
plot(sort(waitingTime), qexp(ppoints(waitingTime), rate = 1 / 32.4), asp = 1, xlab = "Empirical Quantiles", ylab = "Model Quantiles", main = "QQ Plot", col = 'red') # Q-Q plot

abline(a = 0, b = 1) # add the diag line

```

#### **Conclusion: From the Q-Q Plot, we can see that QQ plot deviates most from expected for ISI values between 50-150 ms**.

* * *  

#### 7.Compute the Fano Factor for increments process binned at 1ms, 10ms, 100ms*


```{r, echo = FALSE}
dataforFFin1Ms <- c(rep(1,972), rep(0,30000-972)) # Data for increments process binned at 1ms

dataforFFin10Ms <- c(rep(0,2240), rep(1,574), rep(2,160), rep(3,26)) # Data for increments process binned at 10ms

dataforFFin100Ms <- c(rep(0,22), rep(1,45), rep(2,59), rep(3,61), rep(4,37), rep(5,26), rep(6,23), rep(7,14), rep(8,8), rep(9,3), rep(10,1), rep(11,1)) # Data for increments process binned at 100ms

print(paste("The sample Fano Factor for the increments process binned at 1ms is ", var(dataforFFin1Ms) / mean(dataforFFin1Ms))) # FF 1ms

print(paste("The 95% confidence interval of sample Fano Factor for a Poisson Process at 1ms bin length is [",qgamma(c(0.025,0.975), shape = 15000, rate = 15000)[1],",",qgamma(c(0.025,0.975), shape = 15000, rate = 15000)[2],"]")) # FF 1ms 95% confidence interval 
 
print(paste("The sample Fano Factor for the increments process binned at 10ms is ", var(dataforFFin10Ms) / mean(dataforFFin10Ms))) # FF 10ms

print(paste("The 95% confidence interval of sample Fano Factor for a Poisson Process at 10ms bin length is [",qgamma(c(0.025,0.975), shape = 1500, rate = 1500)[1],",",qgamma(c(0.025,0.975), shape = 1500, rate = 1500)[2],"]")) # FF 10ms 95% confidence interval 

print(paste("The sample Fano Factor for the increments process binned at 100ms is ", var(dataforFFin100Ms) / mean(dataforFFin100Ms))) # FF 100ms

print(paste("The 95% confidence interval of sample Fano Factor for a Poisson Process at 100ms bin length is [",qgamma(c(0.025,0.975), shape = 150, rate = 150)[1],",",qgamma(c(0.025,0.975), shape = 150, rate = 150)[2],"]")) # FF 100ms 95% confidence interval 
```

#### **Conclusion: These statistics do not fall into the 95% confidence interval for a Poisson process at each of these bin lengths**.

* * *

#### 8.Plot the autocorrelation function of the observed interspike intervals with 95% confidence bounds

```{r, echo = FALSE}
plot(acf(waitingTime, lag.max = length(waitingTime), plot = FALSE), type = 'p', pch = 20, col = 'red')
```

#### **Conclusion: From the plot of autocorrelation function, we can see that most of data can be assumed as independent data because there are only a few data falls out the confidence bounds**.

* * *

#### 9.What conclusions can you draw about the spontaneous firing properties of these retinal ganglion neurons & what other variables could be added to improve goodness-of-fit

### **Conclusion: A homogeneous Poisson process fails on many goodness-of-fit tests, so a homogeneous Poisson Process may not be a good choice. The Fano Factor value of 1 ms increments shows that it is more similar to Poisson process, and Fano Factor values of 10 ms and 100 ms increments are significantly greater than 1, which means that comparing to Poisson Process they are less stable. And autocorrelation function shows that data are almost independent, however some of data do not fall into the confidence bounds.**